---
title: "Modeling Crashes with Three Sets of Connectivity Measures"
author: "Jiahua Chen"
date: "2024-01-04"
output: 
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages = c('sf','tidyverse','MASS',"spatialreg","spdep","modelsummary")
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Set up

## Read in

```{r class.source = 'fold-show'}
#newest version: from Dan
var <- st_read("../../scripts/python/outputs/connectivity_blocks.shp")

#turn shp into df
var_df <- var %>%  
  st_drop_geometry()
```

## Clean up

```{r class.source = 'fold-show'}
#common variables
common_var <- var_df %>% 
  dplyr::select(crashes, trips, are_km2 ,slope, pp_dn_2 , phisp, prc_low, incm_md, unmplyd, fem_prc, vtrns_p, wht_prc)
common_var[,-c(1,2)] <- scale(common_var[,-c(1,2)]) #remove those don't need scaling: GEOID, trips, crashes, factor of phisp&income 

#three sets of connectivity measures
connectivity_measures <- list(scale(data.frame(var_df$cnnctvt)),
                              scale(data.frame(var_df$b_net_dens,var_df$b_gamma,var_df$b_cover,var_df$b_int_dens,var_df$b_complex)),
                              scale(data.frame(var_df$ls_net_den,var_df$ls_gamma,var_df$ls_cover,var_df$ls_int_den,var_df$ls_complex)))

```

# Model

## Null Model

```{r}
(model.null <- glm.nb(crashes ~ offset(log(trips)), data = var_df))
```

## Full Model

```{r}
# run models for two different connectivity measure
# set up for loop
# install.packages("rlist")
library(rlist)
model_results_conn <- list()
model_results_all <- list()
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  # model for only offset + connectivity measures
  df <- data.frame(current_measure,"trips"=common_var$trips,"crashes"=common_var$crashes)
  new_model_conn <- glm.nb(crashes ~ . -trips + offset(log(trips)),data = df)
  # model for all variables
  df <- cbind(current_measure,common_var)
  new_model_all <- glm.nb(crashes ~ . -trips + offset(log(trips)),data = df)
  # append model results to the list for making tables
  model_results_conn <- list.append(model_results_conn,new_model_conn)
  model_results_all <- list.append(model_results_all,new_model_all)
}
modelsummary(model_results_conn,stars = TRUE)
modelsummary(model_results_all,stars = TRUE)
```

### Test SAC

All model residuals have a moran'I value at significance level p \> 0.5, so no spatial autocorrelation! Don't need to include spatial error term

```{r}
# test spatial autocorrelation 
# Create a spatial weight matrix
weight_matrix <- poly2nb(var, row.names = var$GEOID)  # Adjust row.names to match your shapefile
weight_listw <- nb2listw(weight_matrix, style = "W")
moran_results <- list()
for (i in seq_along(model_results_all)){
  new_moran <- lm.morantest(model_results_all[[i]],weight_listw)
  moran_results <- list.append(moran_results,new_moran)
  print(new_moran$p.value)
}
```

### Results

For model with only connectivity measures, among the significant coefficients:

1.  Model1-PFB: connectivity score (+)
2.  Model2-bike_network: network density (+), intersection density (-), complexity (-)
3.  Model3-low_stress_network: network density (+), complexity (-)

For model with only all variables, among the significant coefficients:

1.  Model1-PFB: connectivity score (+)
2.  Model2-bike_network: gamma (-)
3.  Model3-low_stress_network: coverage (-)

Notes: for model 2&3, might need to remove percentage_low_stress from the formula as it overlaps with the network coverage variable

## Model with interaction

```{r}
# interaction only for significant connectivity measures
# for now pfb_cnnctvt, b_gamma, ls_cover
model_results_int_sig <- list()
connectivity_measures_sig <- c("var_df.cnnctvt", "var_df.b_gamma", "var_df.ls_cover")
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  df <- cbind(current_measure,common_var)
  formula <- as.formula(paste("crashes ~. -trips + offset(log(trips)) +",
    paste("phisp", connectivity_measures_sig[i], sep = "*")))
  new_model_int <- glm.nb(formula = formula,data = df)
  # append model results to the list for making tables
  model_results_int_sig <- list.append(model_results_int_sig,new_model_int)
}
modelsummary(model_results_int_sig,stars = TRUE)


# loop interaction for all connectivity measures
# set up for looping
model_results_int <- list()
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  df <- cbind(current_measure,common_var)
  formula <- as.formula(paste("crashes ~. -trips + offset(log(trips)) +",
    paste("phisp", colnames(connectivity_measures[[i]]), sep = "*",collapse = "+")))
  new_model_int <- glm.nb(formula = formula,data = df)
  # append model results to the list for making tables
  model_results_int <- list.append(model_results_int,new_model_int)
}
modelsummary(model_results_int,stars = TRUE)

```

### Test SAC

All model residuals have a moran'I value at significance level p \> 0.45, so no spatial autocorrelation! Don't need to include spatial error term

```{r}
# test spatial autocorrelation 
# Create a spatial weight matrix
model_results_int_all <- c(model_results_int,model_results_int_sig)
moran_results_int <- list()
for (i in seq_along(model_results_int_all)){
  new_moran <- lm.morantest(model_results_int_all[[i]],weight_listw)
  moran_results <- list.append(moran_results,new_moran)
  print(new_moran$p.value)
}
# all test p > 0.45, no spatial autocorrelation
```

### Results

For model with interaction of %Hispanic with significant connectivity measures, among the significant coefficients, all interactions are not significant.

For model with interaction of %Hispanic with all connectivity measures, among the significant coefficients:

1.  Model1-PFB: not significant
2.  Model2-bike_network: phisp\*gamma (+), phisp\*coverage(+)
3.  Model3-low_stress_network: phisp\*intersection_density (-)
