---
title: "Modeling Crashes with Three Sets of Connectivity Measures"
author: "Jiahua Chen"
date: "2024-01-04"
output: 
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages = c('sf','tidyverse','MASS',"spatialreg","spdep","modelsummary","labelled")
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Set up

## Read in

```{r class.source = 'fold-show'}
#newest version: with end nodes
var <- st_read("../../scripts/python/outputs/connectivity_blocks/connectivity_blocks.shp")

#turn shp into df
var_df <- var %>%  
  st_drop_geometry()

# turn NAs into 0 for block groups that does not has bike infra present
var_df[is.na(var_df)] <- 0
```

## Clean up

```{r class.source = 'fold-show'}
#common variables
common_var <- var_df %>% 
  dplyr::select(crashes, trips, are_km2 ,slope, pp_dn_2 , phisp, prc_low, incm_md, unmplyd, fem_prc, vtrns_p, wht_prc)
common_var[,-c(1,2)] <- scale(common_var[,-c(1,2)]) #remove those don't need scaling: GEOID, trips, crashes, factor of phisp&income 

#three sets of connectivity measures
connectivity_measures <- list(scale(data.frame(var_df$cnnctvt)),
                              scale(data.frame(var_df$b_net_dens,var_df$b_cover,var_df$b_int_dens,var_df$b_complex)),
                              scale(data.frame(var_df$ls_net_den,var_df$ls_cover,var_df$ls_complex)))




```

# Analysis

## Null Model

```{r}
(model.null <- glm.nb(crashes ~ offset(log(trips)), data = var_df))
model.null.list <- list(model.null)
```

## Model with connectivity indices

```{r}
# run models for two different connectivity measure
# set up for loop
# install.packages("rlist")
library(rlist)

model_results_conn <- list()
model_results_all <- list()
df_list <- list()
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  # model for only offset + connectivity measures
  df <- data.frame(current_measure,"trips"=common_var$trips,"crashes"=common_var$crashes, "slope" =common_var$slope, "area" = common_var$are_km2)
  new_model_conn <- glm.nb(crashes ~ . -trips + offset(log(trips)),data = df)
  # model for all variables
  df <- cbind(current_measure,common_var)
  df_list <- list.append(df_list,df)
  new_model_all <- glm.nb(crashes ~ . -trips -prc_low + offset(log(trips)),data = df)
  # append model results to the list for making tables
  model_results_conn <- list.append(model_results_conn,new_model_conn)
  model_results_all <- list.append(model_results_all,new_model_all)
}
modelsummary(model_results_conn,stars = TRUE)
modelsummary(model_results_all,stars = TRUE)


```

```{r}
list <- c(model_results_conn,model_results_all)
modelsummary(list,
             fmt = fmt_decimal(2, 3), 
             exponentiate = FALSE,
             stars = TRUE,
             coef_rename = c( "slope" = "Slope",
                             "are_km2" = "Area",
                             "area" = "Area"))
```

```{r}
modelsummary(list,
             fmt = fmt_decimal(2, 3), 
             exponentiate = FALSE,
             stars = TRUE,
            coef_rename = c( "slope" = "Slope",
                             "are_km2" = "Area",
                             "area" = "Area",
                             "var_df.cnnctvt" = "Connectivity",
                             "var_df.b_gamma" = "gamma",
                             "var_df.ls_gamma" = "gamma",
                             "var_df.b_net_den" = "Network density",
                             "var_df.b_net_dens" = "Network density",
                             "var_df.ls_net_den" = "Network density",
                             "var_df.b_cover" = "Network coverage",
                             "var_df.ls_cover" = "Network coverage",
                             "var_df.b_int_den" = "Intersections density",
                              "var_df.b_int_dens" = "Intersections density",
                             "var_df.ls_int_den" = "Intersections density",
                             "var_df.b_complex" = "Network complexity",
                             "var_df.ls_complex" = "Network complexity",
                             "phisp" = "% Hispanic",
                             "wht_prc" = "% White",
                             "unmplyd" = "% unemployment",
                             "fem_prc" = "% female",
                             "vtrns_p" = "% veteran",
                             "incm_md" = "Median household income",
                             "pp_dn_2" = "Population density"),
            "result_tables/result_table_nonodes_v4.d")
```

### Test SAC

All model residuals have a moran'I value at significance level p \> 0.5, so no spatial autocorrelation! Don't need to include spatial error term

```{r}
# test spatial autocorrelation 
# Create a spatial weight matrix
weight_matrix <- poly2nb(var, row.names = var$GEOID)  # Adjust row.names to match your shapefile
weight_listw <- nb2listw(weight_matrix, style = "W")
moran_results <- list()
for (i in seq_along(list)){
  new_moran <- lm.morantest(list[[i]],weight_listw)
  moran_results <- list.append(moran_results,new_moran)
  print(new_moran$p.value)
}
```

### Results

For model with only connectivity measures, among the significant coefficients:

1.  Model1-PFB: connectivity score (+)
2.  Model2-bike_network: network density (+), intersection density (-), complexity (-)
3.  Model3-low_stress_network: network density (+), complexity (-)

For model with only all variables, among the significant coefficients:

1.  Model1-PFB: connectivity score (+)
2.  Model2-bike_network: gamma (-)
3.  Model3-low_stress_network: coverage (-)

Notes: for model 2&3, might need to remove percentage_low_stress from the formula as it overlaps with the network coverage variable

## Model with interaction

```{r}
# interaction only for significant connectivity measures
# for now pfb_cnnctvt, b_gamma, ls_cover
model_results_int_sig <- list()
connectivity_measures_sig <- c("var_df.cnnctvt","var_df.ls_net_den","var_df.ls_cover")
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  df <- cbind(current_measure,common_var)
  if (i==2){
    next
  }
  formula <- as.formula(paste("crashes ~. -trips -prc_low + offset(log(trips)) +",
    paste("phisp", connectivity_measures_sig[i], sep = "*")))
  new_model_int <- glm.nb(formula = formula,data = df)
  # append model results to the list for making tables
  model_results_int_sig <- list.append(model_results_int_sig,new_model_int)
}
modelsummary(model_results_int_sig,stars = TRUE)


# loop interaction for all connectivity measures
# set up for looping
model_results_int <- list()
for (i in seq_along(connectivity_measures)) {
  current_measure <- connectivity_measures[[i]]
  df <- cbind(current_measure,common_var)
  formula <- as.formula(paste("crashes ~. -trips -prc_low + offset(log(trips)) +",
    paste("phisp", colnames(connectivity_measures[[i]]), sep = "*",collapse = "+")))
  new_model_int <- glm.nb(formula = formula,data = df)
  # append model results to the list for making tables
  model_results_int <- list.append(model_results_int,new_model_int)
}
modelsummary(model_results_int,stars = TRUE)



```

### Test SAC

All model residuals have a moran'I value at significance level p \> 0.45, so no spatial autocorrelation! Don't need to include spatial error term

```{r}
# test spatial autocorrelation 
# Create a spatial weight matrix
model_results_int_all <- c(model_results_int,model_results_int_sig)
moran_results_int <- list()
for (i in seq_along(model_results_int_all)){
  new_moran <- lm.morantest(model_results_int_all[[i]],weight_listw)
  moran_results <- list.append(moran_results,new_moran)
  print(new_moran$p.value)
}
# all test p > 0.45, no spatial autocorrelation
```

### Results

For model with interaction of %Hispanic with significant connectivity measures, among the significant coefficients, all interactions are not significant.

For model with interaction of %Hispanic with all connectivity measures, among the significant coefficients:

1.  Model1-PFB: not significant
2.  Model2-bike_network: phisp\*gamma (+), phisp\*coverage(+)
3.  Model3-low_stress_network: phisp\*intersection_density (-)

# Output

```{r}
all_model <- c(model.null.list,model_results_conn,model_results_all)

#modelsummary(all_model,stars = TRUE,"result_tables/result_table_withnodes.docx")
modelsummary(all_model,stars = TRUE,"result_tables/result_table_nonodes.docx")
```

## Summary stats

```{r}
connectivity_measures_df <- var_df %>% 
  st_drop_geometry() %>% 
  dplyr::select(cnnctvt, b_gamma , b_cover, b_net_dens ,b_int_dens, b_complex, ls_gamma, ls_cover,ls_net_den, ls_int_den, ls_complex)

datasummary(cnnctvt +  b_gamma + b_cover + b_net_dens + b_int_dens + b_complex + ls_gamma + ls_cover +ls_net_den + ls_int_den + ls_complex ~ Mean + SD + Min +Median +
              Max + Histogram,data = connectivity_measures_df)
```

## Coefficient Plot

```{r}
#install.packages("sjmisc")
#install.packages("sjPlot")
library(sjmisc)
library(sjPlot)

## only connnectivity measure
plot_model(model_results_conn[[1]])
plot_model(model_results_conn[[2]])
plot_model(model_results_conn[[3]])

#with demography
plot_model(model_results_all[[1]],group.terms = c(1,2,2,3,3,3,3,3,3,3))
plot_model(model_results_all[[2]],group.terms = c(1,1,1,1,1,2,2,3,3,3,3,3,3,3))
plot_model(model_results_all[[3]],group.terms = c(1,1,1,1,1,2,2,3,3,3,3,3,3,3))
```

```{r}
plot_model(model_results_all[[1]], type = "eff", terms = "var_df.cnnctvt",data=df_list[[1]])
```

```{r}
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)

# all.models <- list()
# all.models[[1]] <- m1
# all.models[[2]] <- m2
# all.models[[3]] <- m3
model_name <- list("DA_LS","GA_B","GA_LS")


plot_models(model_results_all[[1]],model_results_all[[2]],model_results_all[[3]],
            show.p =TRUE,show.values=TRUE,legend.title = "Connectivity index",
            m.labels = c("Destination Analysis-LS", "Graph Analysis-B", "Graph Analysis-LS"),
            axis.labels = c("complexity-LS","intersection denisty-LS","network coverage-LS","gamma-LS","network density-LS","complexity-B","intersection denisty-B","network coverage-B","gamma-B","network density-B","connectivity"),
             rm.terms = c("are_km2", "slope","pp_dn_2","phisp","incm_md","unmplyd","fem_prc","vtrns_p","wht_prc")) + theme(text = element_text(size = 20))
```

```{r}
library(dplyr)
df <- df_list[[1]] %>% dplyr::select(-prc_low)

test.m <- glm.nb(crashes ~ var_df.cnnctvt + are_km2 + slope + pp_dn_2 + phisp + incm_md + unmplyd + fem_prc + vtrns_p + wht_prc + offset(log(trips)),data = df)

summary(test.m)

library(jtools)
effect_plot(test.m, 
            pred = phisp, 
            data=df,
            interval = TRUE,
            plot.points = TRUE, 
            jitter = 0.05)

```

## Correlation graph

```{r}
cor_sb <- var_df %>% 
  dplyr::select(cnnctvt, b_gamma,b_cover,b_net_dens,b_int_dens,b_complex,ls_gamma,ls_cover,ls_net_den,ls_int_den,ls_complex,pop_dns,phisp,wht_prc,incm_md,unmplyd,fem_prc,vtrns_p) %>% 
  drop_na()

cor_sb <- var_df %>% 
  dplyr::select(cnnctvt, b_gamma,b_cover,b_net_dens,b_int_dens,b_complex,ls_gamma,ls_cover,ls_net_den,ls_int_den,ls_complex) %>% 
  drop_na()

# rename for correlation plot
# colnames(cor_sb) <- c("DA_conn","GA_B_gam","GA_B_cov","GA_B_netD","GA_B_interD","GA_B_comp","GA_LS_gam","GA_LS_cov","GA_LS_netD","GA_LS_interD","GA_LS_comp","pop_dens","perc_hisp","perc_wht","inc_med","unemp","perc_fem","perc_vet")

colnames(cor_sb) <- c("DA_conn","GA_B_gam","GA_B_cov","GA_B_netD","GA_B_interD","GA_B_comp","GA_LS_gam","GA_LS_cov","GA_LS_netD","GA_LS_interD","GA_LS_comp")

# calculate correlation
res <- cor(cor_sb)
```

```{r}
library(corrplot)
## test significance of R2
  # mat : is a matrix of data
  # ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  print(mat)
  n <- ncol(mat)
  print(n)
  p.mat<- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(cor_sb)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(res, method="color", col=col(200),  
         type="lower",
         order="original", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.05, insig = "blank", 
         number.cex=0.5,
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
)
```
